Extracting Function Signature from R Program Traces

### 2020-05-13 paper/main.tex
parametric and higher order fun initial discussion

%
%
%
%
\subsection{Parametric Types and Higher-Order Function Types}

\AT{Ok, high level:
We don't do higher-order function types, beyond any to any.
Typetracer detected them once, but the data was way too noisy.
We found that mostly these correlated with functions where parametric polymorphism would be suitable, though there's no way to say that from the data, and only manual inspection revealed it.
We could implement some heuristic parametric polymorphism detector, but that would be way unsound.
Ultimately, we felt that higher-order function types and parametric polymorphism were related, and they didn't occur frequently enough across all of R.
Some packages are heavy users, but those packages are heavily metaprogrammed and effectively implement DSLs.
We will revisit this issue, but it's not something that we can pick up with \typetracer as it is, and doing justice to these types will require significant effort.}

In R, functions are first-class values, and they can be passed as function arguments.

>>>>

%
%
%
%
\subsection{Parametric and Higher-Order Function Types}

Functions are first-class values in R.

\typetracer found \AT{number} functions to be higher-order, having at least one function type among their arguments.
They are markedly more common in a subset of packages, such as \AT{such as...}.
By-and-large, these higher-order functions are similar to well-known higher-order functions like map, filter, and fold.

The truth is that higher-order function types are best served by {\it parametric types} and parametric polymorphism, wherein type variables are used within a type declaration to specify type-based relationships between function arguments.
Unfortunately, the nature of dynamic analysis is such that \typetracer cannot soundly detect parametric polymorphism, as without a static analysis of the function body it is impossible to say if function execution will halt given certain argument types, and we saw some examples of this in Figures~\ref{fig:scalar-vector-example}~and~\ref{fig:matrix-example}.
Without parametric types, higher-order function types become noisy and fail to tell the complete picture.

Ultimately, we configured \typetracer to report only top-level function types (i.e., detect if a value is a function or not).
\AT{The data suggests that} most of the use of higher-order functions is through map-, filter-, and apply-like functions, whose behaviour is best captured with parametric polymorphism.
Other common uses include higher-order functions operating on data frame columns (in packages such as \code{dplyr}), which would necessitate a more in-depth treatment of data frame types, which as we discussed is a huge endeavour in and of itself.
We aim to extend our type annotation framework with parametric, higher-order, and detailed data frame types together in future work.

%
%
%
%
\subsection{Nullable Types}
